Edge sign prediction in general graphs and Correlation Clustering

### 2017-09-10 troll/introduction.tex
add pointers to other sections in troll intro

When the social network is signed, specific challenges arise in both network analysis and learning.
On the one hand, novel methods are required to tackle standard tasks (e.g., user clustering, link
prediction, targeted advertising/recommendation, prediction of user interests, and analysis of the
spreading of diseases in epidemiological models). On the other hand, new problems such as edge sign
prediction, which we consider here, naturally emerge. Edge sign prediction is the problem of
classifying the positive or negative nature of the links based on the network topology. Prior
knowledge of the network topology is often a realistic assumption, for in several situations the
discovery of the link sign can be more costly than acquiring the topological information of the
network. For instance, when two users of an online social network communicate on a public web page,
we immediately detect a link. Yet, the classification of the link sign as positive or negative may
require complex techniques.

From the modeling and algorithmic viewpoints, because of the huge amount of available networked
data, a major concern in developing learning methods for edge sign prediction is algorithmic
scalability. Many successful, yet simple heuristics for edge sign prediction are based on the
troll-trust features, i.e., on the fraction of outgoing negative links (trollness) and incoming
positive links (trustworthiness) at each node. We study such heuristics by defining a probabilistic
generative model for the signs on the directed links of a given network, and show that these
heuristics can be understood and analyzed as approximators to the Bayes optimal classifier for our
generative model. We also gather empirical evidence supporting our probabilistic model by observing
that a logistic model trained on trollness and trustworthiness features alone is able to learn
weights that, on all datasets considered in our experiments, consistently satisfy the properties
predicted by our model.

We then introduce suitable graph transformations defining reductions from edge sign prediction to
node sign prediction problems. This opens up the possibility of using the arsenal of known
algorithmic techniques developed for node classification. In particular, we show that a Label
Propagation algorithm, combined with our reduction, approximates the maximum likelihood estimator
of our probabilistic generative model. Experiments on real-world data show the competitiveness of
our approach in terms of both prediction performance (especially in the regime when training data
are scarce) and scalability.

Finally, we point out that the notions of trollness and trustworthiness naturally define a measure
of complexity, or learning bias, for the signed network that can also be used to design
\emph{online} (i.e., sequential) learning algorithms for the edge sign prediction problem. The
learning bias encourages settings where the nodes in the network have polarized features (e.g.,
trollness/trustworthiness are either very high or very low). Our online analysis holds under
adversarial conditions, namely, without any stochastic assumption on the assignment of signs to the
network links.

%Hence, it is necessary to depart from well-established yet computationally expensive approaches by relying on novel algorithmic techniques.

%In this paper, we consider the problem of learning link classifiers, and study this problem from different viewpoints. Our contributions are summarized as follows.

>>>>

% When the social network is signed, specific challenges arise in both network analysis and
% learning. On the one hand, novel methods are required to tackle standard tasks (e.g., user
% clustering, link prediction, targeted advertising/recommendation, prediction of user interests,
% and analysis of the spreading of diseases in epidemiological models). On the other hand, new
% problems such as edge sign prediction, which we consider here, naturally emerge.

Such applications motivate studying \esp{}, which is the problem of classifying the positive or
negative nature of the links based on the
network topology. Prior knowledge of the network topology is often a realistic assumption, for in
several situations the discovery of the link sign can be more costly than acquiring the topological
information of the network. For instance, when two users of an online social network communicate on
a public web page, we immediately detect a link. Yet, the classification of the link sign as
positive or negative may require more involved techniques.

From the modeling and algorithmic viewpoints, because of the huge amount of available networked
data, a major concern in developing learning methods for \esp{} is algorithmic
scalability. Many successful, yet simple heuristics for \esp{} are based on the
troll-trust features, \ie{} on the fraction of outgoing negative links (trollness) and incoming
positive links (trustworthiness) at each node. We first define these notions and more notations in
\autoref{s:prel}, along with others tools and precise problem statements. Then we study such
troll-trust heuristics by defining in \autoref{s:gen} a probabilistic
generative model for the signs on the directed links of a given network. We also show that these
heuristics can be understood and analyzed as approximators to the Bayes optimal classifier for our
generative model. In this context, we design our first batch algorithm and show in
\autoref{s:algbatch} that it provably approximates the Bayes classifier on dense graphs.
We furthermore gather empirical evidence supporting our probabilistic model by observing
that a logistic model trained on trollness and trustworthiness features alone is able to learn
weights that, on all datasets considered in our experiments, consistently satisfy the properties
predicted by our model.

We then introduce suitable graph transformations defining reductions from \esp{} to
node sign prediction problems. This opens up the possibility of using the arsenal of known
algorithmic techniques developed for node classification. In particular, we introduce in
\autoref{ss:passive} our second batch algorithm. It takes the form of a Label
Propagation algorithm that, combined with our problem reduction, approximates the maximum likelihood
estimator of our probabilistic generative model.
In order to compare our two algorithms with existing work, we then describe with more details
previous heuristics and related methods in \autoref{sec:troll_related}.
We then experimentally evaluate our proposed approach in \autoref{s:exp}. On synthetic data, we
confirm the quality of our approximation of the Bayes predictor. More importantly, on real-world
data, we show the competitiveness of our approach in terms of both prediction performance
(especially in the regime when training data are scarce) and scalability.

Finally, in \autoref{s:algonline}, we point out that the notions of trollness and trustworthiness
naturally define a measure
of complexity, or learning bias, for the signed network that can also be used to design
\emph{online} (\ie{} sequential) learning algorithms for the \esp{} problem. The
learning bias encourages settings where the nodes in the network have polarized features (e.g.,
trollness/trustworthiness are either very high or very low). Our online analysis holds under
adversarial conditions, namely, without any stochastic assumption on the assignment of signs to the
network links.
